Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-11-18 fib2.tex
holes, 2-curves,..

For our purposes a \emph{surface} will be an
oriented $2$-dimensional manifold with boundary.
We will not require surfaces to be connected.
%Each boundary component is oriented (either explicitly or
%otherwise by using the orientation of the manifold)
Each boundary component will inherit the manifold orientation,
contain a distinguished \emph{base point},
and be labelled with
%an element from a finite set $\mathbb{A}=\{1, a, b, c ...\}.$
an element from a finite set $\A.$
This is the set of ``anyon labels'', and comes
equiped with a vacuum element $\vac$ 
and an involution $\ \widehat{}\ $ such that $\widehat{\vac}=\vac.$

@@

A modular functor $\F$ associates to every
surface $M$ a finite dimensional complex
vector space $\F(M)$, called the \emph{fusion space} of $M$.
A map of surfaces $f:M\to N$ will be understood to be
a topological homeomorphism that preserves boundary
labels, orientation, and base points.
The modular functor associates such an $f$ to a
unitary transformation $\F(f) : \F(M)\to\F(N)$
that only depends on the isotopy class of $f.$
Functoriality requires that $\F$
respect composition of homeomorphisms.

@@

{\bf Axiom III.}
%Gluing.
Denote a surface $M$ with (at least) two boundary
components labelled $a, b$ as $M_{a,b}$. 
Let $N$ be the surface obtained from $M$ by gluing
these two boundary components together.
Glueing must respect the orientation of the two boundary
components as well as identifying their base points.
If we constrain $b=\widehat{a}$ then the fusion space
of $M_{a,\widehat{a}}$ embedds unitarily in the fusion
space of $N.$ Moreover, there is an isomorphism called
a \emph{gluing map}:
$$
    \bigoplus_{a\in\A} \F(M_{a,\widehat{a}}) \xrightarrow{\cong} \F(N).
$$
The seam along which gluing occurs can be associated
with an observable as follows.
We take a gluing map $g$. Projectors $P_a$ onto the
summands in the above direct sum:
$$
    P_a: \bigoplus_{b\in\A} \F(M_{b,\widehat{b}}) \to \F(M_{a,\widehat{a}}).
$$
then the observable will be the set of operators $\{P_a g^{-1}\}_{a\in\A}.$

A common operation is to glue two separate surfaces.
We can do this by first taking disjoint union (tensoring
the fusion spaces)
and then gluing.
Here we show this process applied to two surfaces $M$ and $N$. 
\begin{center}
\includegraphics[]{pic-glue.pdf}
\end{center}
We display the surface $N$ with the $\widehat{a}$
boundary on the outside, 
to show more clearly how $N$ fits into $M$.
In the glued manifold we
indicate the placement of $M$ and $N$ and the seam
along which glueing occured,
as well as the identification of base points.

@@

Here we have shown a manifold $M_{abc}$ with 
three labelled boundaries, as well as two separate
ways of gluing pairs-of-pants to get $M.$

As we have seen, an oriented simple closed path 
 contained in the interior of $M$ can serve as
the seam along which gluing occurs and so we call
these paths \emph{observables}.
%Any oriented simple closed path contained in the interior of $M$ we will
%call an \emph{observable}.
Such a path that encloses two boundary components 
we call a \emph{2-observable}.
\emph{(what does enclose mean exactly?)}
A function $c:[0,1]\to M$ such that $c(0)\in \partial M$ and $c(1)\in \partial M$
are two separate components of the boundary we call a \emph{2-curve.}
The following important fact will be used extensively
in the combinatorial formulation of a modular functor below:
up to isotopy (not preserving $\partial M$),
there is a (natural) bijection between
$2$-observables and $2$-curves.
\emph{(restrict to $M$ a disc with at least two holes?)}
Here we show two such instances of this
bijection:
\begin{center}
\includegraphics[]{pic-2-curve.pdf}
\end{center}
The bijections works as follows.
To construct a $2$-observable from a $2$-curve $c$ that
joins two boundary components $M_1$ and $M_2$,
take a small neighbourhood of
the union $c([0,1])\cup M_1 \cup M_2.$
This neighbourhood will have boundary a simple closed
curve that encloses $M_1$ and $M_2.$
\emph{(check the orientation of the $2$-observable has the correct interior)}
\emph{(the $2$-observable forgets the direction of the $2$-curve)}
Going the other direction...

The end points of $2$-curves are free to terminate anywhere
on the boundary component, but we can fix the endpoints using
the base points.
A simple way to do this is to just require the
$2$-curve to begin and end on base points.

By a \emph{hole} of $M$ we mean a component of $\partial M$.

>>>>

For our purposes a \emph{surface} will be a compact
oriented $2$-dimensional manifold with boundary.
We will not require surfaces to be connected.
%Each boundary component is oriented (either explicitly or
%otherwise by using the orientation of the manifold)
By a \emph{hole} of $M$ we mean a component of its boundary. % $\partial M$.
Each hole will inherit the manifold orientation,
contain a distinguished \emph{base point},
and be labelled with
%an element from a finite set $\mathbb{A}=\{1, a, b, c ...\}.$
an element from a (fixed) finite set $\A.$
This is the set of ``anyon labels'', and comes
equiped with a vacuum element $\vac$ 
and an involution $\ \widehat{}\ $ such that $\widehat{\vac}=\vac.$

@@

A map of surfaces $f:M\to N$ will be understood to be
a topological homeomorphism that preserves
manifold orientation, hole labels, and base points.

A modular functor $\F$ associates to every
surface $M$ a finite dimensional complex
vector space $\F(M)$, called the \emph{fusion space} of $M$.
For each map $f:M\to N$
The modular functor associates a
unitary transformation $\F(f) : \F(M)\to\F(N)$
that only depends on the isotopy class of $f.$
Functoriality requires that $\F$
respect composition of maps.

@@

{\bf Axiom III.}
%Gluing.
Denote a surface $M$ with (at least) two holes
labelled $a, b$ as $M_{a,b}$. 
Let $N$ be the surface obtained from $M$ by gluing
these two holes together.
Gluing must respect the orientation of the two holes
as well as identifying their base points.
If we constrain $b=\widehat{a}$ then the fusion space
of $M_{a,\widehat{a}}$ embedds unitarily in the fusion
space of $N.$ Moreover, there is an isomorphism called
a \emph{gluing map}:
$$
    \bigoplus_{a\in\A} \F(M_{a,\widehat{a}}) \xrightarrow{\cong} \F(N).
$$
The seam along which gluing occurs can be associated
with an observable as follows.
We take a gluing map $g$. Projectors $P_a$ onto the
summands in the above direct sum:
$$
    P_a: \bigoplus_{b\in\A} \F(M_{b,\widehat{b}}) \to \F(M_{a,\widehat{a}}).
$$
then the observable will be the set of operators $\{P_a g^{-1}\}_{a\in\A}.$

A common operation is to glue two separate surfaces.
We can do this by first taking disjoint union (tensoring
the fusion spaces)
and then gluing.
Here we show this process applied to two surfaces $M$ and $N$. 
\begin{center}
\includegraphics[]{pic-glue.pdf}
\end{center}
We display the surface $N$ with the $\widehat{a}$
boundary on the outside, 
to show more clearly how $N$ fits into $M$.
In the glued manifold we
indicate the placement of $M$ and $N$ and the seam
along which gluing occured,
as well as the identification of base points.

@@

Here we have shown a manifold $M_{abc}$ with 
three labelled boundaries, as well as two separate
ways of gluing pairs-of-pants to get $M.$
\emph{XXX define pair-of-pants}

As we have seen, an oriented simple closed path 
contained in the interior of $M$ can serve as
the seam along which gluing occurs and we call
these paths \emph{observables}.
%Any oriented simple closed path contained in the interior of $M$ we will
%call an \emph{observable}.
Such a path that encloses two holes
we call a \emph{2-observable}.
\emph{(what does enclose mean exactly?)}
A function $c:[0,1]\to M$ such that 
$c(0)\in \partial M$ and $c(1)\in \partial M$
are distinct basepoints (of two holes)
we call a \emph{2-curve.}
%The following important fact will be used extensively
%in the combinatorial formulation of a modular functor below:
Up to isotopy, % (not preserving $\partial M$),
there is a (natural) bijection between
$2$-observables and $2$-curves.
\emph{(XXX we really need undirected $2$-curves)}
%\emph{(restrict to $M$ a disc with at least two holes?)}
Here we show two such instances of this
bijection:
\begin{center}
\includegraphics[]{pic-2-curve.pdf}
\end{center}
The bijections works as follows.
To construct a $2$-observable from a $2$-curve $c$ that
joins two holes $M_1$ and $M_2$,
take a small neighbourhood of
the union $c([0,1])\cup M_1 \cup M_2.$
This neighbourhood will have boundary a simple closed
curve that encloses $M_1$ and $M_2.$
\emph{(is this really how to prove this? do we care?)}
\emph{(check the orientation of the $2$-observable has the correct interior)}
Going the other direction we can take
any $2$-curve contained within the interior of a $2$-observable,
\emph{but how do we show these are all isotopic?}

%The end points of $2$-curves are free to terminate anywhere
%on the hole, but we can fix the endpoints using
%the base points.
%A simple way to do this is to just require the
%$2$-curve to begin and end on base points.
